What can GitHub and Googlecode Issues tell us about Exceptions in Android?

### 2014-08-27 android-stacks.tex
changesl all over

Untill recent years, the availability of issues reporting abnormal
computation states (e.g., crashes) was scarce. The
available data comprised only few large scale projects (e.g. the Mozilla
crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Recently we have observed a rapid increase in the number of open
source projects that make their control version as well as the issue tracking systems
publicly available in repository hosting sites such as GitHub and Googlecode.

@@

When available, crash-related information added on issues and crash reports can enable different kinds of post-mortem analysis, such as
 bug classification and clustering ~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  the automated bug fixing ~\cite{sinha2009fault},  
fault-proneness prediction models ~\cite{kim2013predicting} and can also be used to support developers in debugging ~\cite{schroter2010stack}.

Most of such studies, however, focused the analysis on one or two system at a time. None 
of them target crash-related information of family of Android applications 
(which share a common core) aiming at  extracting their common characteristics 
 and looking for similar vulnerabilities on applications and underlaying platform.

In this work, we perform the first post mortem analysis of the exception stack traces embedded on the issues
of all Android applications available in GitHub and Googlecode. Using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from issues across all Android projects on
GitHub and Googlecode. Overall, we analyzed the XXXX issues extracted from on XXX Android projects defined in
GitHub and XXX projects defined in Googlecode. From these issues  XXX stack traces were extracted and used as
 a source of information for our analysis. The stack trace analysis was augmented by additional information
extracted from JDK, Android SO and Android SDK using byte code and source code analysis - which was done 
on a carefully selected subset of our initial sample (482 Android projects). Manual inspection was
used to leverage the understanding of exception stack traces and support further discussions and insights.

@@

This study is built on the following assumption: the stack traces reported on issues contain 
relevant information related to application crashes. Several studies and techniques have also based on this assumption (e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}).

@@

Falar da assumption que stacj trá relacionada a crash?

@@

@@

\subsection{Direct instances of RuntimeExceptions being thrown.}

@@

\subsubsection{Classifying Top 100 Exceptions}

@@

@@

To answer this question we: (i) distilled the information available on stack traces (i.e., exception signalers, root causes, exception wrappings); 
(ii) mined the exception types related to each exception reported on stack traces using source code and byte code analyses; and finally checked 
this information against the compiled best practices. Doing so, we mined some characteristics of exception stack traces that points to best practices violations.

@@

\subsection{What can the Exception Wrappings tell us?}

Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
stack traces found in Android repositories. As we can see, some of the checked
exceptions were indeed wrapped in runtime exceptions or even errors, while most
of them were not wrapped along the stack trace.

@@

Such wrapping was detected on distinct 70 projects. Inspecting the 
elements responsible for performing such wrapping we could observe that,
most of such wrappings were perform on methods defined on Android.

\begin{table}
\centering
\begin{tabular}{ll}
    \hline
    \bfseries{signaler from} & \bfseries{Ocurrences} \\
    \hline
ANDROID	 & 66 \\
APP	& 29 	(HOW MANY?? AND WHICH??) \\
LIB  & 19 (HOW MANY AND WHICH?)\\
\hline
  \end{tabular}
\caption{Elements responsible for the wrapping.}
\label{tab:wrapping01}
\end{table}

Taking a closer look on the signalers reposponsible for wrapping we could observe that, aproximately 41%
of all wrappings were performed by a single class, called ActivityThread, responsible for managing the life cycle 
of an Android application. It performs the application callbacks.

\begin{table}
\centering
\begin{tabular}{ll}
 \bfseries{Signaler class} & \bfseries{Ocurrences} \\
    \hline
android.app.ActivityThread & 47 \\
android.view.LayoutInflater & 7 \\
org.gege.caldavsyncadapter.syncadapter.SyncAdapter & 7 \\
org.springframework.web.client.RestTemplate & 6 \\
android.os.AsyncTask & 5 \\
\hline
  \end{tabular}
\caption{Elements tresponsible for the wrapping.}
\label{tab:wrapping01}
\end{table}

Code snipet extracted from android.app.ActivityThread.

@@

private Activity performLaunchActivity(...) {
  try {
   ...
  } catch (Exception e) {
    ...
    throw new RuntimeException(..., e);
  }
}
\end{verbatim}
}

The exceptions most wrapped on RuntimeExceptinos were:

\begin{table}
\centering
\begin{tabular}{lll}
    \hline
    \bfseries{wrapper} & \bfseries{root} & \bfseries{\#} \\
    \hline
java.lang.RuntimeException & java.lang.ClassNotFoundException &	32 \\
java.lang.RuntimeException & java.lang.NoSuchMethodException &	6 \\
java.lang.RuntimeException & java.lang.InstantiationException) & 5 \\
...andlyticsproject.NetworkException & org.apache...HttpResponseException &	5\\
android.view.InflateException & java.lang.NoSuchMethodException & 4 \\
\hline
  \end{tabular}
\caption{Elements tresponsible for the wrapping.}
\label{tab:wrapping01}
\end{table}

\begin{table}
\centering
\begin{tabular}{lll}
    \hline
    \bfseries{Exception} & \bfseries{Ocurrences} \\
    \hline
java.lang.RuntimeException &	69 & 61\% \\
android.view.InflateException	& 10 & 8,7\% \\
com.github.andlyticsproject.console.NetworkException	& 7 & - \\
java.lang.IllegalStateException	& 6 & - \\

\hline
  \end{tabular}
\caption{Elements tresponsible for the wrapping.}
\label{tab:wrapping01}
\end{table}

@@

\subsection{Undocumented runtime exceptions.} After filtering out all the exceptions implicitly 
signaled by  JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an indirect 
instances of java.lang.RuntimeException). We could observe that only 1 method out of 118 inspected signalers
 (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
throws clause on method signature). This result is aligned with the results of other  study conducted by from 
Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
runtime exceptions in .Net programs are most often not documented.

@@

%\noindent\emph{The high number of null pointer exceptions}

@@

@@

\noindent\emph{External Validity} The results presented here were based on mining
issues on Github through the GHTorrent dataset. While comprehensive and
extensive, GHTorrent is not an exact replica of Github, so several issues might
be left out. Due to the way data collection works with GHTorrent, for projects
that are relatively inactive, GHTorrent might have not collected a significant
proportion of their issues. We did not investigate the extend of this threat on
our sample.

>>>>

In Java, when an application fails due to an uncaught exception, 
it automatically terminates, while the system prints a stack trace to the console, or log file [REF-STACK-TRACE-CRASH]. 
A stack trace includes a detail message and the execution stack frame when 
the exception occurred. A typical Java stack trace consists of an ordered list of 
methods that were active on the call stack before the exception has occurred [REF-STACKTRACE-MANUAL].

Untill recent years, the availability of issues reporting abnormal
computation states (e.g., exception stack traces) was scarce. The
available data comprised only few large scale projects (e.g. the Mozilla
crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Recently we have observed a rapid increase in the number of open
source projects that make their control version as well as the issue tracking systems
publicly available in repository hosting sites such as GitHub and Googlecode.

@@

%Stack traces are a useful source of information about system crashes, and is often used to support developers in debugging~\cite{schroter2010stack}. Moreover, information %from stack traces combined with crash and bug reports can be used to support bug classification and clustering~\cite{wang2013improving, kim2011crash, %dhaliwal2011classifying}, fault-proneness prediction models~\cite{kim2013predicting} and even automated bug fixing~\cite{sinha2009fault} tools.

When available, exception stack traces provides a useful source of information about system crashes and 
 can enable different kinds of post-mortem analysis, such as:   debugging~\cite{schroter2010stack}, 
 bug classification and clustering ~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  
automated bug fixing ~\cite{sinha2009fault} and fault-proneness prediction models analysis ~\cite{kim2013predicting}.

Most of such studies, however, focused the analysis on one or two system at a time. None 
of them targeted a family of applications (sharing a common framework) aiming at extracting 
the common characteristics of crash information and checking whether they could provide
insights on  similar vulnerabilities on the applications and underlaying framework.

In this work, we perform the first post mortem analysis of the exception stack traces embedded on all issues
created on Android projects available in GitHub and Googlecode. Using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from issues across all Android projects on
GitHub and Googlecode. Overall, we analyzed the XXXX issues extracted from on XXX Android projects defined in
GitHub and XXX projects defined in Googlecode. From these issues,  XXX stack traces were extracted and used as
 a source of information for our analysis. The stack trace analysis was augmented by additional information
extracted from JDK, Android SO and Android SDK and the applications using byte code and source code analysis.
 Manual inspection was used to leverage the understanding of exception stack traces and support further discussions and insights.

@@

@@

@@

Falar da assumption que stacj trá relacionada a crash?
This study is built on the following assumption: the stack traces reported on issues contain 
relevant information related to application crashes. Several studies and techniques have also based on this assumption (e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}).

@@

\subsubsection{Direct instances of RuntimeExceptions being thrown.}

@@

\subsection{Classifying Top 100 Exceptions}

@@

\subsection{Characterizing the Most Common Root Exceptions Signalers}
<COMPLEMENT.... INCLUDE THE METHODS AND CLASSES RESPONSIBLE FOR SIGNALING
MOST OF THE EXCEPTIONS AND THE EXCEPTIONS SIGNALED BY THEM. SIMILAR TO THE WRAPPINGS TABLES>

@@

To answer this question, besides distilling the information available on stack traces (i.e., exception signalers, root causes, exception wrappings),
we mined the exception types (i.e., RuntimeException, checked Exception or Error) related to each exception reported on stack traces 
using source code and byte code analyses supported by Exception Miner; and finally checked 
this information against the compiled best practices. Doing so, we mined some characteristics 
of exception stack traces that points to best practices violations.

@@

\subsection{What the Types of Root Exceptions can tell us?}
\subsection{Exploring the Information Embeeded on Exception Types}

<falar dos tipos mais comuns.... Runtime e NullPointer???>
<LINKCAR COM ANALISE DOS SUPERTIPOS>
<complementar falando do grafico acima>
\subsubsection{Undocumented runtime exceptions.} After filtering out all the exceptions implicitly 
signaled by  JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an indirect 
instances of java.lang.RuntimeException). We could observe that only 1 method out of 118 inspected signalers
 (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
throws clause on method signature). This result is aligned with the results of other  study conducted by from 
Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
runtime exceptions in .Net programs are most often not documented.

<fazer o link com wrappings>

\subsection{What the Exception Wrappings can Tell us?}
\subsection{Exploring the Information Embeeded on Exception Wrappings}
Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
stack traces found in Android repositories. As we can see, some of the checked
exceptions were indeed wrapped in runtime exceptions or even errors, while most
of them were not wrapped along the stack trace.

@@

Such wrapping was detected on 70 distinct projects. Inspecting the 
elements responsible for performing such wrapping, we could observe that
most of such wrappings were performed on methods defined on Android
Platform (approximately 58%).

%\begin{table}
%\centering
%\begin{tabular}{ll}
%    \hline
%   \bfseries{signaler from} & \bfseries{Ocurrences} \\
%    \hline
%ANDROID	 & 66 \\
%APP	& 29 	(HOW MANY?? AND WHICH??) \\
%LIB  & 19 (HOW MANY AND WHICH?)\\
%\hline
% \end{tabular}
%\caption{Elements responsible for the wrapping.}
%\label{tab:wrapping01}
%\end{table}

Taking a closer look on the methods reposponsible for such wrappings,
 we could observe that aproximately 41% of them were performed by a single class,
 named ActivityThreat. This class is responsible for managing the life cycle of an Android Activities.

\begin{table}
\centering
\begin{tabular}{lll}
 \bfseries{Signaler class} &  \bfseries{Wrapper Type} & \bfseries{Ocurrences} \\
    \hline
android.app.ActivityThread & java.lang.RuntimeException & 47 \\
android.view.LayoutInflater & android.view.InflateException &  7 \\
android.os.AsyncTask & java.lang.RuntimeException &  5 \\
Other platform classes & java.lang.RuntimeException &   7 \\
\hline
  \end{tabular}
\caption{Elements tresponsible for the wrapping.}
\label{tab:wrapping01}
\end{table}

Figure X presents a code snippet of ActivityThreat which is reponsible for (among other things)
calling the Activities onCreate method. Hence, as we can see from lines 4-8,
 in case any exception is thrown by an application Activity during its creation, 
it is wrapped in a general Runtime exception and rethrown.

@@

1. private Activity performLaunchActivity(...) {
2.  try {
3.   ...
4.   ...callActivityOnCreate(...);
5.  } catch (Exception e) {
6.    ...
7.   throw new RuntimeException(..., e);
8.  }
9.}
\end{verbatim}
}

The inversion of control makes it dificult to the application developer to associate 
a specific handler to the exceptions thrown during the onCreate method (or any other Activity life cycle method).
Since according to Java handling model the handler should me added on any method in the call chain,
and in the inversion of control development model the developer usually does not have fully access to 
the invocation call chain of a method in order to associate a handler, as a result the 
exception may scape from any handling and remain uncaught.

Difficulty of associating a specific handler to exceptions signaled by application methods called by the framework,
may also be the reason behind the design desicion of encapsulatling any exception in a RuntimeException.
As framework does not have enought contextual information the exception regardless its types,
it encapsulates in a general type and re-throws. The only possible answer to such exceptions would be 
to finish the application.

Such design decision may lead the developers into believing that by just re-throwing 
the exceptions they can forget about the exceptional situations during the development of the "happy path". 
This "ignore-for-now" approach may turn the exception handling into a generalized "goto" mechanism~\cite{mandrioli1992advances}
 making the program more complex and even less reliable. This behaviour may lead to the well known uncaught exceptions 
problem~\cite{jo2004uncaught}, one of the main causes of application crashes.

% Several types of exceptions were  wrapped the RuntimeExceptinos thrown by the 
% ActivityThread. Table X illustrates them.
% #\begin{table}
% \centering
% \begin{tabular}{lll}
%    \hline
%    \bfseries{wrapper} & \bfseries{root} & \bfseries{\#} \\
%    \hline
% java.lang.RuntimeException & java.lang.ClassNotFoundException &	32 \\
% java.lang.RuntimeException & java.lang.NoSuchMethodException &	6 \\
% java.lang.RuntimeException & java.lang.InstantiationException) & 5 \\
% ...andlyticsproject.NetworkException & org.apache...HttpResponseException &	5\\
% android.view.InflateException & java.lang.NoSuchMethodException & 4 \\
% \hline
%  \end{tabular}
% \caption{Elements tresponsible for the wrapping.}
% \label{tab:wrapping01}
% \end{table}

% \begin{table}
% \centering
% \begin{tabular}{lll}
%   \hline
%   \bfseries{Exception} & \bfseries{Ocurrences} \\
%   \hline
% java.lang.RuntimeException &	69 & 61\% \\
% android.view.InflateException	& 10 & 8,7\% \\
% com.github.andlyticsproject.console.NetworkException	& 7 & - \\
% java.lang.IllegalStateException	& 6 & - \\
% \hline
%  \end{tabular}
% \caption{Elements tresponsible for the wrapping.}
% \label{tab:wrapping01}
% \end{table}

@@

@@

%\noindent\emph{The high number of NullPointerExceptions and other JVM-signaled Exceptions}

Moreover, some languages use exceptions to represent coding errors (out-of-bounds array index, division-by-zero, access to a null). In these cases exceptions are implicitly signaled by the runtime environment when such condition happens. One of such implicit exceptions (caused by data conversion from a 64-bit floating point to a 16-bit signed integer) was the responsible for the famous failure on Ariane 5's first test flight~\cite{lions1996ariane} --- leading to the rocket self-destructing 37 seconds after launch and a loss of 500 million dollars. Besides Ariane 5, several applications crash everyday due to uncaught exceptions~\cite{jo2004uncaught}. The consequences of not careful exception usage may have the opposite effect of its initial intention of improving system robustness.

@@

%\noindent\emph{Unwritten Rules of Android Exception Handling Developement}

@@

\noindent\emph{External Validity} The results presented here were based on mining
issues on Github through the GHTorrent dataset. While comprehensive and
extensive, GHTorrent is not an exact replica of Github, so several issues might
be left out. Due to the way data collection works with GHTorrent, for projects
that are relatively inactive, GHTorrent might have not collected a significant
proportion of their issues. We did not investigate the extend of this threat on
our sample.
<OLHAR ARTIGO DO FSE QUE TRABALHA COM OPENSOURCE.
FALAR QUE DESCOBERTAS FORAM DE LIBS JAVA. APESAR DO ESTUDO NAO SER FEITO EM PROJS JAVA.....
PODEMOS CONCLUIR ALGUMAS COISAS QUE SE APLICAM A TODOS JAVA>
